---
title: "ddpart"
output:
 pdf_document:
  toc: true
  toc_depth: 3
vignette: >
 %\VignetteIndexEntry{ddpart}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

\newpage

# Introduction

This package is an implementation of the particle dry deposition model of Zhang
et al. (2001) including the parametrization from Emerson et al. (2020). This
vignette contains an example application of the package and validation against
the original publication (Emerson et al. (2020)).

Usually, the wrapper functions CalculateDepositionVelocity() or
CalculateDepositionVelocity2() would be used. These functions include the
calculation of atmospheric stability conditions (Monin-Obukhov-Length) and
friction velocity via the Pasquill class approach. This has the advantage that
widely available meteorological measurements (or modelled data) can be used as
an input (e.g. wind speed, solar radiation, etc.). If this is not required
because direct data on friction velocity is available, the underlying functions
to calculate the resistance components can be called directly to skip the
Pasquill class approach. See section "Validation" for examples where this is the
case.

# Package installation and dependencies

Install the package:
```{r install, message = FALSE, results = 'hide'}
devtools::install_github(
  repo = "https://github.com/AndSchmitz/ddpart"
)
library(ddpart)
```

Dependencies for this vignette:
```{r deps}
library(tidyr)
library(knitr)
library(scales)
library(ggplot2)
```

\newpage

# Usage example

The following example shows how to use the package to calculate the dry
deposition velocity of particles with 5 um diameter to grassland for a 48 h time
series of meteorological data.

## Load and show example meteo time series
```{r show meteo time series, fig.width=8, fig.height=10, fig.cap = "Example time series of meteorological data."}
meteo_time_series_long <- meteo_time_series %>%
  pivot_longer(
    cols = -c(TimeStampUTC)
  )

ggplot(
  data = meteo_time_series_long,
  mapping = aes(
    x = TimeStampUTC,
    y = value
  )
) +
  geom_line() +
  facet_wrap(~name, scales = "free_y", ncol = 3) +
  scale_x_datetime(guide = guide_axis(n.dodge = 2))
```


## Calculate dry deposition velocity

In the following code block, the function CalculateDepositionVelocity()
is used. This function requires a dataframe as input ("InputTable"). For each
row of InputTable, a deposition velocity is calculated. The InputTable must
contain all information required to calculate aerodynamic resistance and 
surface resistance (e.g. roughness length, zero plane distance height,
wind speed and anemomemter height, hight of measured (or modelled)
concentration data (reference height), etc.). See ?CalculateDepositionVelocity
for a description.

The function CalculateDepositionVelocity() is designed for the usual setup where
both, wind speed measurements and concentrations measurements, are available
over the same land use (e.g. both measured over grassland). For other cases,
see ?CalculateDepositionVelocity2. For example, if wind speed data refers to
grassland conditions (e.g. from meteorological models) but the dry deposition
velocity should be calculated for other land uses (e.g. forest).


```{r calculate vd, fig.width=8, fig.height=8, fig.cap = "Dry deposition velocity calculated for the 48 hours of input data."}

InputTable <- meteo_time_series %>%
  mutate(
    LUCNames = "Grassland",
    RoughnessLength_m = 0.03,
    ZeroPlaneDisplacementHeight_m = 7 * RoughnessLength_m,
    ReferenceHeight_m = 20, # Height of concentration measurements
    DryParticleDiameter_m = 5 * 1e-6, # 5 um particles
    ParticleDensity_kgm3 = 2000,
    AerosolType = "SeaSalt", # Relevant for hygroscopic swelling
    Parametrization = "Zhang01",
    SurfaceIsVegetated_bool = T
  )


Results <- CalculateDepositionVelocity(
  InputTable = InputTable
)


ResultsLong <- Results %>%
  mutate(
    DryDepositionVelocity_cms = V_d_RefHeight_ms * 100
  ) %>%
  select(TimeStampUTC, DryDepositionVelocity_cms) %>%
  pivot_longer(
    cols = -c(TimeStampUTC)
  )

ggplot(
  data = ResultsLong,
  mapping = aes(
    x = TimeStampUTC,
    y = value
  )
) +
  geom_line() +
  facet_wrap(~name, scales = "free_y", nrow = 3) +
  scale_x_datetime(guide = guide_axis(n.dodge = 2))
```


\newpage

# Validation

Emerson et al. (2020) figures 1 and 2 show the dry deposition velocity as a a
function of particle diameter for different settings (i.e. for different land
use types, friction velocity, etc.). In the following, these figures are
reproduced as a validation of the functions in the ddpart package. Emerson et
al. (2020) used a specific implementation of the Zhang01 approach in the
GEOS-Chem model (https://github.com/geoschem/geos-chem/). This implementation
differs from the original formulation by Zhang01 in some aspects. In total,
there are 4 variants of parametrizations:

  A) "Zhang01": The original model according to Zhang et al. (2001). Not used
  for the figures in Emerson et al. (2020). Implemented in ddpart as
  parametrization "Zhang01".

  B) "GEOS-Chem old": This is what Emerson et al. (2020) refer to as "Zhang et
  al. (2001)" in their Fig. 1 and 2. Implemented in ddpart as
  parametrization "GCOld". It is an adaption of parametrization A) with two
  differences:

  - First, the land use classes in the GEOS-Chem model differ from those in A).
  The mapping between land use classes is not always straightforward. For
  example, "Grassland" in B) uses parameters for "shrubs and interrupted
  woodland" from A). The complete mapping of LUCs is defined in file
  "drydep_mod.F90" starting in lines 3143:
  https://github.com/geoschem/geos-chem/blob/main/GeosCore/drydep_mod.F90.

  - Second, no differentiation according to seasons exists. Instead, the average
  over seasons 1-5 is used for each parameter. See for example
  "drydep_mod.F90" line 3226.

  C)  "GEOS-Chem new": This is the "revised parametrization" in Emerson et al.
  Fig. 1 and 2. Implemented in ddpart as parametrization "GCNew". It uses the
  same assignment of LUCs as in B) and also averages land use parameters over
  seasons, but some  parameters have been re-calibrated by Emerson et al. to
  better match measurement data. Note that although Emerson applies the
  parameters for "Shrubs and interrupted wood-lands" (LUC10 from Zhang01) for
  grassland, the parametrization is valid for grassland. This is because
  Emerson20 *calibrated* the other parameters, such that the resulting dry
  deposition velocity matches measurement data for grassland.

  D) "GCNewSeason": Same as C) but with season-specific land use parameters.  E.g.
  parameter "characteristic receptor radius" (A in mm) varies between seasons
  for deciduous broadleaf forest. This is implemented in ddpart as
  parametrization "GCNewSeason".

In order to reproduce Fig. 1 and Fig. 2 from Emerson et al. (2020), variants B)
and C) must be used.

\newpage

## 1. Contribution from dry deposition sub-processes

Emerson et al. (2020) figure 2 shows the contribution of the four dry deposition
sub-processes (Brownian motion, interception, impaction and gravitational
settling) to the total dry deposition velocity. This figure is reconstructed
with functions from the ddpart package. 

### Definition of parameters

```{r Emerson fig 2 parameters}

# Land use parameters for variants B) and C)
LUCPars <- expand_grid(
  LUCNames = "Needleleaf",
  Season = 1, # irrelevant for GCOld and GCNew parametrizations
  Parametrization = c("GCOld", "GCNew") #see explanation text above
) %>%
  mutate(
    A_mm = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "A_mm"
    ),
    alpha = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "alpha"
    ),
    gamma = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "gamma"
    )
  )

# Vector of particle diameters
d_p_vec <- 10^seq(-2, 2, length.out = 5)
tmp <- expand.grid(1:9, d_p_vec)
d_p_vec <- tmp[, 1] * tmp[, 2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec <- d_p_vec * 1e-6

# Other parameters
OtherPars <- data.frame(
  R_a_sm = 0,
  ParticleDensity_kgm3 = 1500,
  FrictionVelocity_ms = 0.4,
  SurfaceIsVegetated_bool = T
) %>%
  mutate(
    # These parameters are guessed
    # Standard temperatur and pressure
    T_air_K = 293.15,
    AirPressure_Pa = 101325,
    SurfaceIsWet_bool = F
  )

# Repeat input rows for each particle diameter
Input <- bind_cols(LUCPars, OtherPars) %>%
  expand_grid(
    ParticleDiameter_m = d_p_vec
  )
```

\newpage

### Calculations

```{r Emerson fig 2B Calculations}
Output <- Input %>%
  mutate(
    # Meteo basics
    DynamicViscosityAir_kgms = CalculateDynamicViscosityOfAir(
      T_air_K = T_air_K
    ),
    AirDensity_kgm3 = CalculateAirDensity(
      AirPressure_Pa,
      T_air_K
    ),
    KinematicViscosityOfAir_m2s = CalculateKinematicViscosityOfAir(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      AirDensity_kgm3 = AirDensity_kgm3
    ),
    MeanFreePathOfAirMolecule_m = CalculateMeanFreePath(
      T_air_K = T_air_K,
      AirPressure_Pa = AirPressure_Pa,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # Deposition processes------
    SettlingVelocity_ms = CalculateSettlingVelocity(
      ParticleDensity_kgm3 = ParticleDensity_kgm3,
      ParticleDiameter_m = ParticleDiameter_m,
      MeanFreePathOfAirMolecule_m = MeanFreePathOfAirMolecule_m,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # _Schmidt number------
    SchmidtNumber = CalculateSchmidtNumber(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      T_air_K = T_air_K,
      ParticleDiameter_m = ParticleDiameter_m
    ),
    # _E_b-----
    # Loss efficiency by Brownian diffusion
    BrownianDiffusionParameterGamma = gamma,
    E_b = CalculateLossEfficiencyBrownianDiffusion(
      SchmidtNumber = SchmidtNumber,
      BrownianDiffusionParameterGamma = BrownianDiffusionParameterGamma,
      Parametrization = Parametrization
    ),
    # _Stokes number-----
    CharacteristicRadius_m = A_mm / 1000,
    StokesNumber = CalculateStokesNumber(
      FrictionVelocity_ms = FrictionVelocity_ms,
      SettlingVelocity_ms = SettlingVelocity_ms,
      CharacteristicRadius_m = CharacteristicRadius_m,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      SurfaceIsVegetated = SurfaceIsVegetated_bool
    ),
    # _E_Im----
    # Loss efficiency by impaction
    ImpactionParameterAlpha = alpha,
    E_Im = CalculateLossEfficiencyImpaction(
      StokesNumber = StokesNumber,
      ImpactionParameterAlpha = ImpactionParameterAlpha,
      Parametrization = Parametrization
    ),
    # _E_In----
    # Loss efficiency by interception
    E_In = CalculateLossEfficiencyInterception(
      ParticleDiameter_m = ParticleDiameter_m,
      CharacteristicRadius_m = CharacteristicRadius_m,
      Parametrization = Parametrization
    ),
    # _R_s-----
    # Surface resistance
    R_s_sm = CalculateSurfaceResistance(
      SurfaceIsWet = SurfaceIsWet_bool,
      FrictionVelocity_ms = FrictionVelocity_ms,
      StokesNumber = StokesNumber,
      E_b = E_b,
      E_Im = E_Im,
      E_In = E_In,
      ParticleDiameter_m = ParticleDiameter_m,
      Parametrization = Parametrization
    ),
    # _V_d---------
    V_d_s = 1 / (R_a_sm + R_s_sm),
    V_g = SettlingVelocity_ms,
    V_d_ms = V_d_s + V_g,
    Type = "Results from ddpart"
  )

# Plot separate dry deposition processes contributions
ProcessContributions <- Output %>%
  mutate(
    BounceCorrectionTerm = ifelse(
      test = SurfaceIsWet_bool,
      yes = 1,
      no = exp(-sqrt(StokesNumber))
    ),
    # Calculation of resistances and corresponding vd's for each process
    epsilon_0 = GetParameters(TargetParameter = "epsilon_0", Parametrization = Parametrization),
    R_s_E_b_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_b),
    V_d_E_b_only = 1 / (R_a_sm + R_s_E_b_only),
    R_s_E_Im_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_Im),
    V_d_E_Im_only = 1 / (R_a_sm + R_s_E_Im_only),
    R_s_E_In_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_In),
    V_d_E_In_only = 1 / (R_a_sm + R_s_E_In_only)
  ) %>%
  select(Parametrization, ParticleDiameter_m, V_d_E_b_only, V_d_E_Im_only, V_d_E_In_only, V_d_ms, SettlingVelocity_ms) %>%
  pivot_longer(
    cols = -c("Parametrization", "ParticleDiameter_m"),
    names_to = "Component",
    values_to = "v"
  ) %>%
  mutate(
    ComponentName = case_when(
      Component == "V_d_E_b_only" ~ "Brownian",
      Component == "V_d_E_Im_only" ~ "Impaction",
      Component == "V_d_E_In_only" ~ "Interception",
      Component == "SettlingVelocity_ms" ~ "Settling",
      Component == "V_d_ms" ~ "Total"
    )
  )
```

\newpage

### Summary of parameters

```{r Emerson fig 2 Pars}

TableData <- Output %>%
  select(
    Parametrization, LUCNames,
    T_air_K, AirPressure_Pa, SurfaceIsWet_bool,
    SurfaceIsVegetated_bool, CharacteristicRadius_m,
    ImpactionParameterAlpha, ParticleDensity_kgm3,
    BrownianDiffusionParameterGamma,
    R_a_sm
  ) %>%
  distinct() %>%
  rename(
    Gamma = BrownianDiffusionParameterGamma
  ) %>%
  mutate(
    Gamma = round(Gamma, 3),
    C_b = GetParameters(Parametrization = Parametrization, TargetParameter = "C_b"),
    nu = GetParameters(Parametrization = Parametrization, TargetParameter = "nu"),
    C_In = GetParameters(Parametrization = Parametrization, TargetParameter = "C_In"),
    beta = GetParameters(Parametrization = Parametrization, TargetParameter = "beta"),
    C_Im = GetParameters(Parametrization = Parametrization, TargetParameter = "C_Im")
  ) %>%
  pivot_longer(
    cols = -c("Parametrization"),
    values_transform = as.character,
    names_to = "Parameter"
  ) %>%
  pivot_wider(
    names_from = "Parametrization",
    values_from = "value"
  )

kable(
  x = TableData
)
```

### Plots

```{r Emerson fig 2 Plots, fig.width=8, fig.height=5, fig.cap = "Contribution of Brownian diffusion, Gravitational settling, Interception and Impaction to the total dry deposition velocity. Lines indicate data extracted from fig. 2 in Emerson et al. (2020). Results from the ddpart R package are indicated by +. The Left panel shows parametrization 'GEOS-Chem old' which is similar to the Zhang et al. (2001) parametrization. The right panel shows parametrization 'GEOS-Chem new' which is referred to as the 'revised' parametrization in Emerson et al. (2020). The figures show results at 0.4 m/s friction velocity for needleleaf forest and a particle density of 1500 kg/m3. Data from ddpart is generated with aerodynamic resistance set to zero, no hygroscopic swelling and dry surface. See parameter table above for full information."}

# Load data extracted from Emerson et al. (2020)
ValidationData <- dd_subprocess_validation %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("GCOld", "GCNew")
    ),
    Type = "Extracted from Emerson et al. (2020)"
  )
ProcessContributions <- ProcessContributions %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("GCOld", "GCNew")
    ),
    Type = "Calculated with ddpart"
  )

Colors <- c("black", "blue", "orange", "darkred", "purple")
names(Colors) <- c("Total", "Brownian", "Settling", "Interception", "Impaction")
Shapes <- c(3)
names(Shapes) <- c("Calculated with ddpart")
LineTypes <- c("solid")
names(LineTypes) <- c("Extracted from Emerson et al. (2020)")

ggplot() +
  geom_line(
    data = ValidationData,
    mapping = aes(
      x = ParticleDiameter_um,
      y = DepositionVelocity_cms,
      color = Process,
      linetype = Type
    )
  ) +
  geom_point(
    data = ProcessContributions,
    mapping = aes(
      x = ParticleDiameter_m * 1e6,
      y = v * 100,
      color = ComponentName,
      shape = Type
    ),
    size = 2
  ) +
  scale_shape_manual(values = Shapes) +
  scale_color_manual(values = Colors) +
  scale_linetype_manual(values = LineTypes) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3, 1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.box = "vertical",
    panel.spacing = unit(1, "lines")
    # legend.margin=margin()
  ) +
  facet_wrap(~Label)
```

\newpage

## 2. Dry deposition velocity vs. particle size for 3 different land use classes

Emerson et al. (2020) figure 1 shows the dry deposition velocity for different
land use classes. Note that Emerson et al. state that their Fig. 1 has been
generated with a friction velocity of 0.2 m/s. Three observations strongly
suggest that this is not the case and that instead the same friction velocity as
in Emerson et al. Fig. 2 (0.4 m/s) has been used by Emerson to generate Fig. 1.
These observations are:

- The relation between dry deposition velocity and particle size for needleleaf
forest in Emerson Fig. 1 is *exactly* identical to Emerson Fig. 2. This is shown
below based on the data extracted from the Emerson paper. Emerson Fig. 2 has
been generated with friction velocity of 0.4 m/s according to its figure
caption.

- Dry deposition velocity is directly dependend on friction velocity (e.g. via
surface resistance, see Emerson eq. 2). I.e. a change of friction velocity from
0.4 m/s to 0.2 m/s would create obvious differences in the relation between dry
deposition velocity and particle size (which is not seen in Emerson Fig. 1 vs.
Emerson Fig. 2).

Therefore, a friction velocity of 0.4 m/s is used to reproduce Emerson Fig. 1.

### Definition of parameters

```{r Emerson fig 1 parameters}

# LUC pars
LUCPars <- expand_grid(
  LUCNames = c("Grassland", "Needleleaf", "DecBroadleaf"),
  Season = 1, # irrelevant for GCOld and GCNew parameterizations
  Parametrization = c("GCOld", "GCNew") #see explanation text above
) %>%
  mutate(
    A_mm = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "A_mm"
    ),
    alpha = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "alpha"
    ),
    gamma = GetLandUseParameters(
      LUCNames = LUCNames,
      Seasons = Season,
      Parametrizations = Parametrization,
      TargetPar = "gamma"
    )
  )

# Vector of particle diameters
d_p_vec <- 10^seq(-2, 2, length.out = 5)
tmp <- expand.grid(1:9, d_p_vec)
d_p_vec <- tmp[, 1] * tmp[, 2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec <- d_p_vec * 1e-6

# Other parameters
OtherPars <- data.frame(
  # See above why FrictionVelocity_ms is 0.4 and not 0.2 m/s as stated in caption
  # of Emerson20 Fig1.
  FrictionVelocity_ms = 0.4,
  ParticleDensity_kgm3 = 1200,
  SurfaceIsVegetated = TRUE,
  # These parameters are guessed:
  # "article water uptake corresponding to ambient relative humidity"
  RelHum_percent = 80,
  AerosolType = "Rural",
  # Standard temperature and pressure
  T_air_K = 293.15,
  AirPressure_Pa = 101325,
  R_a_sm = 0,
  SurfaceIsWet_bool = FALSE
)

# Repeat input rows for each particle diameter
Input <- bind_cols(LUCPars, OtherPars) %>%
  expand_grid(
    DryParticleDiameter_m = d_p_vec
  )
```

\newpage

### Calculations

```{r Emerson fig 1 Calculations}
Output <- Input %>%
  mutate(
    ParticleDiameter_m = CalculateHygroscopicSwelling(
      DryParticleDiameter_m = DryParticleDiameter_m,
      RelHum_percent = RelHum_percent,
      AerosolType = AerosolType
    ),
    # Meteo basics
    DynamicViscosityAir_kgms = CalculateDynamicViscosityOfAir(
      T_air_K = T_air_K
    ),
    AirDensity_kgm3 = CalculateAirDensity(
      AirPressure_Pa,
      T_air_K
    ),
    KinematicViscosityOfAir_m2s = CalculateKinematicViscosityOfAir(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      AirDensity_kgm3 = AirDensity_kgm3
    ),
    MeanFreePathOfAirMolecule_m = CalculateMeanFreePath(
      T_air_K = T_air_K,
      AirPressure_Pa = AirPressure_Pa,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # Deposition processes------
    SettlingVelocity_ms = CalculateSettlingVelocity(
      ParticleDensity_kgm3 = ParticleDensity_kgm3,
      ParticleDiameter_m = ParticleDiameter_m,
      MeanFreePathOfAirMolecule_m = MeanFreePathOfAirMolecule_m,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # _Schmidt number------
    SchmidtNumber = CalculateSchmidtNumber(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      T_air_K = T_air_K,
      ParticleDiameter_m = ParticleDiameter_m
    ),
    # _E_b-----
    E_b = CalculateLossEfficiencyBrownianDiffusion(
      SchmidtNumber = SchmidtNumber,
      BrownianDiffusionParameterGamma = gamma,
      Parametrization = Parametrization
    ),
    StokesNumber = CalculateStokesNumber(
      FrictionVelocity_ms = FrictionVelocity_ms,
      SettlingVelocity_ms = SettlingVelocity_ms,
      CharacteristicRadius_m = A_mm / 1000,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      SurfaceIsVegetated = SurfaceIsVegetated
    ),
    # _E_Im----
    # Loss efficiency by impaction
    E_Im = CalculateLossEfficiencyImpaction(
      StokesNumber = StokesNumber,
      ImpactionParameterAlpha = alpha,
      Parametrization = Parametrization
    ),
    # _E_In----
    # Loss efficiency by interception
    E_In = CalculateLossEfficiencyInterception(
      ParticleDiameter_m = ParticleDiameter_m,
      CharacteristicRadius_m = A_mm / 1000,
      Parametrization = Parametrization
    ),
    # _R_s-----
    # Surface resistance
    R_s_sm = CalculateSurfaceResistance(
      SurfaceIsWet = SurfaceIsWet_bool,
      FrictionVelocity_ms = FrictionVelocity_ms,
      StokesNumber = StokesNumber,
      E_b = E_b,
      E_Im = E_Im,
      E_In = E_In,
      ParticleDiameter_m = ParticleDiameter_m,
      Parametrization = Parametrization
    ),
    # _V_d---------
    V_d_surface_ms = 1 / (R_a_sm + R_s_sm),
    V_g = SettlingVelocity_ms,
    V_d_ms = V_d_surface_ms + V_g,
    Type = "Results from ddpart"
  )
```

\newpage

### Summary of parameters

```{r Emerson fig 1 Pars}

TableData <- Output %>%
  select(
    Parametrization, LUCNames,
    T_air_K, AirPressure_Pa, SurfaceIsWet_bool,
    SurfaceIsVegetated, A_mm,
    alpha, ParticleDensity_kgm3,
    gamma, R_a_sm
  ) %>%
  distinct() %>%
  mutate(
    gamma = round(gamma, 3),
    C_b = GetParameters(Parametrization = Parametrization, TargetParameter = "C_b"),
    nu = GetParameters(Parametrization = Parametrization, TargetParameter = "nu"),
    C_In = GetParameters(Parametrization = Parametrization, TargetParameter = "C_In"),
    beta = GetParameters(Parametrization = Parametrization, TargetParameter = "beta"),
    C_Im = GetParameters(Parametrization = Parametrization, TargetParameter = "C_Im")
  ) %>%
  pivot_longer(
    cols = -c("Parametrization", "LUCNames"),
    values_transform = as.character,
    names_to = "Parameter"
  ) %>%
  mutate(
    LUCShort = case_when(
      LUCNames == "Needleleaf" ~ "NL",
      LUCNames == "DecBroadleaf" ~ "BL",
      LUCNames == "Grassland" ~ "GR",
      T ~ "ERROR"
    ),
    Parametrization_LUC = paste0(Parametrization, "_", LUCShort)
  ) %>%
  select(-Parametrization, -LUCNames, -LUCShort) %>%
  pivot_wider(
    names_from = "Parametrization_LUC",
    values_from = "value"
  )

kable(
  x = TableData
)
```

\newpage

### Plots

```{r Emerson fig 1 Plots, fig.width=8, fig.height=4, fig.cap="Dry deposition velocity as a function of particle diameter for needleleaf forest, decidous broadleaf forest and grassland. Lines represent data extracted from fig. 1 in Emerson et al. (2020). Results from the ddpart R package are indicated by +. The figures show results at 0.4 m/s friction velocity and a particle density of 1200 kg/m3. Data from ddpart is generated with aerodynamic resistance set to zero, dry surface, 80% relative humidity and corresponding hygroscopic swelling for aerosol of type 'rural' (Zhang et al. (2001) eq. 10). Particle diameter refers to the dry particle diameter (i.e. before accounting for hygroscopic swelling)."}

# Load data extracted from Emerson et al. (2020)
ValidationData <- dd_validation %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("GCOld", "GCNew")
    ),
    Type = "Extracted from Emerson et al. (2020)",
    LUCLabel = factor(
      x = LUCName,
      levels = c("Needleleaf", "DecBroadleaf", "Grassland")
    )
  )
Output <- Output %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("GCOld", "GCNew")
    ),
    Type = "Calculated with ddpart",
    LUCLabel = factor(
      x = LUCNames,
      levels = c("Needleleaf", "DecBroadleaf", "Grassland")
    )
  )

Shapes <- c(3)
names(Shapes) <- c("Calculated with ddpart")
LineTypes <- c("solid")
names(LineTypes) <- c("Extracted from Emerson et al. (2020)")

ggplot() +
  geom_line(
    data = ValidationData,
    mapping = aes(
      x = ParticleDiameter_um,
      y = DepositionVelocity_cms,
      color = Parametrization,
      linetype = Type
    )
  ) +
  geom_point(
    data = Output,
    mapping = aes(
      x = DryParticleDiameter_m * 1e6,
      y = V_d_ms * 100,
      color = Parametrization,
      shape = Type
    ),
    size = 2
  ) +
  scale_shape_manual(values = Shapes) +
  scale_linetype_manual(values = LineTypes) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3, 1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.box = "horizontal",
    panel.spacing = unit(1, "lines")
    # legend.margin=margin()
  ) +
  facet_wrap(~LUCLabel)
```


\newpage


### Emerson Fig 1. vs. Emerson Fig. 2

The following graph shows that the relation between dry deposition velocity and
particle size are exactly identical in Emerson Fig. 1 and Emerson Fig. 2. This
strongly suggests that Emerson used the same value of friction velocity to
generate these graphs.

 - Variable "dd_validation" contains the modelled dry deposition velocity for
 needleleaf forest, deciduous broadleaf forest and grassland extracted from
 Emerson Fig 1. Here, only the data for needleleaf forest is used.
 
 - Variable "dd_subprocess_validation" contains the modelled dry deposition
 velocity for needleleaf forest extracted from Emerson Fig. 2 ("total dry
 deposition" is used).
 
 - In both cases, data has been extracted using a plot digitalizer tool. 

 
```{r Ecmp, fig.width=8, fig.height=4, fig.cap = "Modelled dry deposition velocity extracted from Emerson Fig. 1 and Fig. 2 is identical."}
EmersonFig1 <- dd_validation %>%
  filter(
    LUCName == "Needleleaf"
  ) %>%
  select(-LUCName) %>%
  mutate(
    DataSource = "Emerson et al. (2020) Fig1"
  )

EmersonFig2 <- dd_subprocess_validation %>%
  filter(
    Process == "Total"
  ) %>%
  select(-Process) %>%
  mutate(
    DataSource = "Emerson et al. (2020) Fig2"
  )

EmersonCmp <- bind_rows(EmersonFig1, EmersonFig2) %>%
  mutate(
    Parametrization = factor(
      x = paste0("Parametrization = ", Parametrization),
      levels = c("Parametrization = GCOld", "Parametrization = GCNew")
    )
  )

ggplot(
  data = EmersonCmp,
  mapping = aes(
    x = ParticleDiameter_um,
    y = DepositionVelocity_cms,
    color = DataSource,
    linetype = DataSource
  )
) +
  geom_line(
    size = 2
  ) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3, 1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  facet_wrap(~Parametrization) +
  theme(
    legend.position = "bottom"
  )
```


\newpage

# References

 - Seinfeld JH, Pandis SN. Atmospheric Chemistry and Physics: From Air Pollution to Climate Change. Second edition. Wiley; 2006.

 - Zhang L, Gong S, Padro J, Barrie L. A size-segregated particle dry deposition scheme for an atmospheric aerosol module. Atmospheric Environment 2001;35:549–560.

 - Emerson EW, Hodshire AL, DeBolt HM, Bilsback KR, Pierce JR, McMeeking GR, Farmer DK. Revisiting particle dry deposition and its role in radiative effect estimates. Proceedings of the National Academy of Sciences 2020;117:26076–26082.
