---
title: "ddpart"
author: "Andreas Schmitz"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{ddpart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Package installation

Install the package
```{r install, message = FALSE}
devtools::install_github(
  repo = "https://github.com/AndSchmitz/ddpart"
)
library(ddpart)
```

Dependencies for this vignette
```{r deps}
library(tidyr)
library(knitr)
library(scales)
library(ggplot2)
```

# Usage example

## Load and show example meteo time series
```{r show meteo time series, fig.width=8, fig.height=12}
meteo_time_series_long <- meteo_time_series %>%
  pivot_longer(
    cols = -c(TimeStampUTC)
  )

ggplot(
  data = meteo_time_series_long,
  mapping = aes(
    x = TimeStampUTC,
    y = value
  )
) +
geom_line() +
facet_wrap( ~name, scales = "free_y", ncol = 3) +
  scale_x_datetime(guide = guide_axis(n.dodge = 2))
```

## Calculate dry deposition velocity
```{r calculate vd, fig.width=8, fig.height=8}


#Final settings. RoughnessLength_m and ZeroPlaneDisplacementHeight_m must be
#identical for wind speed measurements and for concentrations measurements
#(i.e. both have been measured/modelled over the same land use type). If this
#is not the case, use function CalculateDepositionVelocity2().
InputTable <- meteo_time_series %>%
  mutate(
    LUCZhang2001 = 6, #grassland
    RoughnessLength_m = 0.03,
    ZeroPlaneDisplacementHeight_m = 7 * RoughnessLength_m,
    ReferenceHeight_m = 20, #Height of concentration measurements
    DryParticleDiameter_m = 5 * 1e-6, #5 um particles
    ParticleDensity_kgm3 = 2000,
    AerosolType = "SeaSalt",
    Parametrization = "Emerson20"
  )


Results <- CalculateDepositionVelocity(
  InputTable = InputTable
)


ResultsLong <- Results %>%
  mutate(
    DryDepositionVelocity_cms = V_d_RefHeight_ms * 100,
    WetParticleDiameter_um = ParticleDiameter_m * 1e6
  ) %>%
  select(TimeStampUTC, DryDepositionVelocity_cms, WetParticleDiameter_um, WindSpeedAtAnemometerHeight_ms) %>%
  pivot_longer(
    cols = -c(TimeStampUTC)
  )

ggplot(
  data = ResultsLong,
  mapping = aes(
    x = TimeStampUTC,
    y = value
  )
) +
  geom_line() +
  facet_wrap( ~name, scales = "free_y", nrow = 3) +
  scale_x_datetime(guide = guide_axis(n.dodge = 2))
```

\newpage

# Validation

Emerson et al. (2020) figures 1 and 2 show the dry deposition velocity as a a
function of particle diameter for different settings (i.e. for different land
use types, friction velocity, etc.). In the following, these figures are
reproduced as a validation of the functions in the ddpart package.

## 1. Dry deposition velocity vs. particle size

Emerson et al. (2020) figure 1 shows the dry deposition velocity for different
land use classes. This figure is reconstructed with functions from the ddpart
package. 


### Definition of parameters

```{r Emerson fig 1 parameters}

# Vector of particle diameters
d_p_vec <- 10^seq(-2, 2, length.out = 5)
tmp <- expand.grid(1:9, d_p_vec)
d_p_vec <- tmp[, 1] * tmp[, 2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec <- d_p_vec * 1e-6

InputPars <- data.frame(
  R_a_sm = 0,
  Parametrization = "Emerson20",
  LUC = 1,
  Season = 1,
  ParticleDensity_kgm3 = 1200,
  FrictionVelocity_ms = 0.2,
  AerosolType = "Rural",
  RelHum_percent = 85,
  # These parameters are guessed
  # Standard temperature and pressure
  T_air_K = 293.15,
  AirPressure_Pa = 101325,
  SurfaceIsWet_bool = TRUE  
)
  
# Duplicate input for Zhang01 parametrization
tmp <- InputPars %>%
  mutate(
    Parametrization = "Zhang01"
  )
InputPars <- bind_rows(InputPars, tmp)

# Duplicate input for broadleaf and grassland
tmp1 <- InputPars %>%
  mutate(
    LUC = 2
  )
tmp2 <- InputPars %>%
  mutate(
    LUC = 6
  )
InputPars <- bind_rows(InputPars, tmp1, tmp2)

Input <- expand_grid(
  InputPars,
  DryParticleDiameter_m = d_p_vec
) %>%
  mutate(
    ParticleDiameter_m = CalculateHygroscopicSwelling(
      DryParticleDiameter_m = DryParticleDiameter_m,
      RelHum_percent = RelHum_percent,
      AerosolType = AerosolType
    )
  )

```

### Calculations

```{r Emerson fig 1 Calculations}
Output <- Input %>%
  mutate(
    # Meteo basics
    DynamicViscosityAir_kgms = CalculateDynamicViscosityOfAir(
      T_air_K = T_air_K
    ),
    AirDensity_kgm3 = CalculateAirDensity(
      AirPressure_Pa,
      T_air_K
    ),
    KinematicViscosityOfAir_m2s = CalculateKinematicViscosityOfAir(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      AirDensity_kgm3 = AirDensity_kgm3
    ),
    MeanFreePathOfAirMolecule_m = CalculateMeanFreePath(
      T_air_K = T_air_K,
      AirPressure_Pa = AirPressure_Pa,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # Deposition processes------
    SettlingVelocity_ms = CalculateSettlingVelocity(
      ParticleDensity_kgm3 = ParticleDensity_kgm3,
      ParticleDiameter_m = ParticleDiameter_m,
      MeanFreePathOfAirMolecule_m = MeanFreePathOfAirMolecule_m,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # _Schmidt number------
    SchmidtNumber = CalculateSchmidtNumber(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      T_air_K = T_air_K,
      ParticleDiameter_m = ParticleDiameter_m
    ),
    # _E_b-----
    # Loss efficiency by Brownian diffusion
    BrownianDiffusionParameterGamma = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "gamma",
      Parametrization = Parametrization
    ),
    E_b = CalculateLossEfficiencyBrownianDiffusion(
      SchmidtNumber = SchmidtNumber,
      BrownianDiffusionParameterGamma = BrownianDiffusionParameterGamma,
      Parametrization = Parametrization
    ),
    # _Stokes number-----
    SurfaceIsVegetated = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "SurfaceIsVegetated",
      Parametrization = Parametrization
    ),
    CharacteristicRadius_m = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "A_mm",
      Parametrization = Parametrization
    ) * 1e-3,
    StokesNumber = CalculateStokesNumber(
      FrictionVelocity_ms = FrictionVelocity_ms,
      SettlingVelocity_ms = SettlingVelocity_ms,
      CharacteristicRadius_m = CharacteristicRadius_m,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      SurfaceIsVegetated = SurfaceIsVegetated
    ),
    # _E_Im----
    # Loss efficiency by impaction
    ImpactionParameterAlpha = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "alpha",
      Parametrization = Parametrization
    ),
    E_Im = CalculateLossEfficiencyImpaction(
      StokesNumber = StokesNumber,
      ImpactionParameterAlpha = ImpactionParameterAlpha,
      Parametrization = Parametrization
    ),
    # _E_In----
    # Loss efficiency by interception
    E_In = CalculateLossEfficiencyInterception(
      ParticleDiameter_m = ParticleDiameter_m,
      CharacteristicRadius_m = CharacteristicRadius_m,
      Parametrization = Parametrization
    ),
    # _R_s-----
    # Surface resistance
    R_s_sm = CalculateSurfaceResistance(
      SurfaceIsWet = SurfaceIsWet_bool,
      FrictionVelocity_ms = FrictionVelocity_ms,
      StokesNumber = StokesNumber,
      E_b = E_b,
      E_Im = E_Im,
      E_In = E_In,
      Parametrization = Parametrization
    ),
    # _V_d---------
    V_d_s = 1 / (R_a_sm + R_s_sm),
    V_g = SettlingVelocity_ms,
    V_d_ms = V_d_s + V_g,
    Type = "Results from ddpart"
  )
```

\newpage

### Parameter table

```{r Emerson fig 1 Pars}

TableData <- Output %>%
  select(
    Parametrization, LUC, Season,
    T_air_K, AirPressure_Pa, SurfaceIsWet_bool,
    SurfaceIsVegetated, CharacteristicRadius_m,
    ImpactionParameterAlpha, ParticleDensity_kgm3,
    BrownianDiffusionParameterGamma,
    R_a_sm
  ) %>%
  distinct() %>%
  rename(
    Gamma = BrownianDiffusionParameterGamma
  ) %>%
    mutate(
      Gamma = round(Gamma,3),
    C_b = GetParameters(Parametrization = Parametrization, TargetParameter = "C_b"),
    nu = GetParameters(Parametrization = Parametrization, TargetParameter = "nu"),
    C_In = GetParameters(Parametrization = Parametrization, TargetParameter = "C_In"),
    beta = GetParameters(Parametrization = Parametrization, TargetParameter = "beta"),
    C_Im = GetParameters(Parametrization = Parametrization, TargetParameter = "C_Im")
  ) %>%
  pivot_longer(
    cols = -c("Parametrization", "LUC"),
    values_transform = as.character,
    names_to = "Parameter"
  ) %>%
  mutate(
    Parametrization_LUC = paste0(Parametrization, "_LUC", LUC)
  ) %>%
  select(-Parametrization, -LUC) %>%
  pivot_wider(
    names_from = "Parametrization_LUC",
    values_from = "value"
  )

kable(
  x = TableData
)
```

\newpage

### Plots

FIXME - these deviations might be explained by some differences in the
formulations of hygroscopic growth or the slip correction factors between
Emerson et al. (2020) using the GEOS-CHEM model and the R implementation using
hygroscopic growth from Zhang et al. (2001) and the slip correction from
Seinfeld and Pandis (2006).

```{r Emerson fig 1 Plots, fig.width=8, fig.height=4}
# Load data extracted from Emerson et al. (2020)
ValidationData <- dd_validation %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("Zhang01", "Emerson20")
    ),
    Type = "Extracted from Emerson et al. (2020)",
    LUCLabel = case_when(
      LUC == 1 ~ "Needleleaf forest",
      LUC == 2 ~ "Broadlead forest",
      LUC == 6 ~ "Grassland",
      T ~ "ERROR"
    ),
    LUCLabel = factor(
      x = LUCLabel,
      levels = c("Needleleaf forest", "Broadlead forest", "Grassland")
    )
  )
Output <- Output %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("Zhang01", "Emerson20")
    ),
    Type = "Calculated with ddpart",
    LUCLabel = case_when(
      LUC == 1 ~ "Needleleaf forest",
      LUC == 2 ~ "Broadlead forest",
      LUC == 6 ~ "Grassland",
      T ~ "ERROR"
    ),
    LUCLabel = factor(
      x = LUCLabel,
      levels = c("Needleleaf forest", "Broadlead forest", "Grassland")
    )    
  )

Shapes <- c(3)
names(Shapes) <- c("Calculated with ddpart")
LineTypes <- c("solid")
names(LineTypes) <- c("Extracted from Emerson et al. (2020)")

ggplot() +
  geom_line(
    data = ValidationData,
    mapping = aes(
      x = ParticleDiameter_um,
      y = DepositionVelocity_cms,
      color = Parametrization,
      linetype = Type
    )
  ) +
  geom_point(
    data = Output,
    mapping = aes(
      x = DryParticleDiameter_m * 1e6,
      y = V_d_ms * 100,
      color = Parametrization,
      shape = Type
    ),
    size = 2
  ) +
  scale_shape_manual(values = Shapes) +
  scale_linetype_manual(values = LineTypes) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3, 1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.box = "horizontal",
    panel.spacing = unit(1, "lines")
    # legend.margin=margin()
  ) +
  facet_wrap(~LUCLabel)

```
Figure 1: Dry deposition velocity as a function of particle diameter for needleleaf forest (LUC 1), evergreen broadleaf forest (LUC 2) and grassland (LUC 6) according to the Zhang et al. (2001) parametrization and the Emerson et al. (2020) parameterization. Lines represent data extracted from fig. 1 in Emerson et al. (2020). Results from the ddpart R package are indicated by "+". The figures show results at 0.2 m/s friction velocity and a particle density of 1200 kg/m3. Data from ddpart is generated for season 1 with aerodynamic resistance set to zero, wet surface, 85% relative humidity and corresponding hygroscopic swelling for aerosol of type "rural" (Zhang et al. (2001) eq. 10). Particle diameter refers to the dry particle diameter (i.e. before accounting for hygroscopic swelling).


## 2. Contribution from dry deposition sub-processes

Emerson et al. (2020) figure 2 shows the contribution of the four dry deposition
sub-processes (Brownian motion, interception, impaction and gravitational
settling) to the total dry deposition velocity. This figure is reconstructed
with functions from the ddpart package. 

### Definition of parameters

```{r Emerson fig 2 parameters}

# Vector of particle diameters
d_p_vec <- 10^seq(-2, 2, length.out = 5)
tmp <- expand.grid(1:9, d_p_vec)
d_p_vec <- tmp[, 1] * tmp[, 2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec <- d_p_vec * 1e-6

InputPars <- data.frame(
  R_a_sm = 0,
  Parametrization = "Emerson20",
  LUC = 1,
  Season = 1,
  ParticleDensity_kgm3 = 1500,
  FrictionVelocity_ms = 0.4
) %>%
  mutate(
    # These parameters are guessed
    # Standard temperatur and pressure
    T_air_K = 293.15,
    AirPressure_Pa = 101325,
    SurfaceIsWet_bool = F
  )

# Duplicate input for Zhang01 parametrization
tmp <- InputPars %>%
  mutate(
    Parametrization = "Zhang01"
  )
InputPars <- bind_rows(InputPars, tmp)

Input <- expand_grid(
  InputPars,
  ParticleDiameter_m = d_p_vec
)

```

### Calculations

```{r Emerson fig 2B Calculations}
Output <- Input %>%
  mutate(
    # Meteo basics
    DynamicViscosityAir_kgms = CalculateDynamicViscosityOfAir(
      T_air_K = T_air_K
    ),
    AirDensity_kgm3 = CalculateAirDensity(
      AirPressure_Pa,
      T_air_K
    ),
    KinematicViscosityOfAir_m2s = CalculateKinematicViscosityOfAir(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      AirDensity_kgm3 = AirDensity_kgm3
    ),
    MeanFreePathOfAirMolecule_m = CalculateMeanFreePath(
      T_air_K = T_air_K,
      AirPressure_Pa = AirPressure_Pa,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # Deposition processes------
    SettlingVelocity_ms = CalculateSettlingVelocity(
      ParticleDensity_kgm3 = ParticleDensity_kgm3,
      ParticleDiameter_m = ParticleDiameter_m,
      MeanFreePathOfAirMolecule_m = MeanFreePathOfAirMolecule_m,
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms
    ),
    # _Schmidt number------
    SchmidtNumber = CalculateSchmidtNumber(
      DynamicViscosityAir_kgms = DynamicViscosityAir_kgms,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      T_air_K = T_air_K,
      ParticleDiameter_m = ParticleDiameter_m
    ),
    # _E_b-----
    # Loss efficiency by Brownian diffusion
    BrownianDiffusionParameterGamma = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "gamma",
      Parametrization = Parametrization
    ),
    E_b = CalculateLossEfficiencyBrownianDiffusion(
      SchmidtNumber = SchmidtNumber,
      BrownianDiffusionParameterGamma = BrownianDiffusionParameterGamma,
      Parametrization = Parametrization
    ),
    # _Stokes number-----
    SurfaceIsVegetated = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "SurfaceIsVegetated",
      Parametrization = Parametrization
    ),
    CharacteristicRadius_m = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "A_mm",
      Parametrization = Parametrization
    ) * 1e-3,
    StokesNumber = CalculateStokesNumber(
      FrictionVelocity_ms = FrictionVelocity_ms,
      SettlingVelocity_ms = SettlingVelocity_ms,
      CharacteristicRadius_m = CharacteristicRadius_m,
      KinematicViscosityOfAir_m2s = KinematicViscosityOfAir_m2s,
      SurfaceIsVegetated = SurfaceIsVegetated
    ),
    # _E_Im----
    # Loss efficiency by impaction
    ImpactionParameterAlpha = GetLandUseParameters(
      LUCs = LUC,
      Seasons = Season,
      TargetPar = "alpha",
      Parametrization = Parametrization
    ),
    E_Im = CalculateLossEfficiencyImpaction(
      StokesNumber = StokesNumber,
      ImpactionParameterAlpha = ImpactionParameterAlpha,
      Parametrization = Parametrization
    ),
    # _E_In----
    # Loss efficiency by interception
    E_In = CalculateLossEfficiencyInterception(
      ParticleDiameter_m = ParticleDiameter_m,
      CharacteristicRadius_m = CharacteristicRadius_m,
      Parametrization = Parametrization
    ),
    # _R_s-----
    # Surface resistance
    R_s_sm = CalculateSurfaceResistance(
      SurfaceIsWet = SurfaceIsWet_bool,
      FrictionVelocity_ms = FrictionVelocity_ms,
      StokesNumber = StokesNumber,
      E_b = E_b,
      E_Im = E_Im,
      E_In = E_In,
      Parametrization = Parametrization
    ),
    # _V_d---------
    V_d_s = 1 / (R_a_sm + R_s_sm),
    V_g = SettlingVelocity_ms,
    V_d_ms = V_d_s + V_g,
    Type = "Results from ddpart"
  )

# Plot separate dry deposition processes contributions
ProcessContributions <- Output %>%
  mutate(
    BounceCorrectionTerm = ifelse(
      test = SurfaceIsWet_bool,
      yes = 1,
      no = exp(-sqrt(StokesNumber))
    ),
    # Calculation of resistances and corresponding vd's for each process
    epsilon_0 = GetParameters(TargetParameter = "epsilon_0", Parametrization = Parametrization),
    R_s_E_b_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_b),
    V_d_E_b_only = 1 / (R_a_sm + R_s_E_b_only),
    R_s_E_Im_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_Im),
    V_d_E_Im_only = 1 / (R_a_sm + R_s_E_Im_only),
    R_s_E_In_only = 1 / (epsilon_0 * FrictionVelocity_ms * BounceCorrectionTerm * E_In),
    V_d_E_In_only = 1 / (R_a_sm + R_s_E_In_only)
  ) %>%
  select(Parametrization, ParticleDiameter_m, V_d_E_b_only, V_d_E_Im_only, V_d_E_In_only, V_d_ms, SettlingVelocity_ms) %>%
  pivot_longer(
    cols = -c("Parametrization", "ParticleDiameter_m"),
    names_to = "Component",
    values_to = "v"
  ) %>%
  mutate(
    ComponentName = case_when(
      Component == "V_d_E_b_only" ~ "Brownian",
      Component == "V_d_E_Im_only" ~ "Impaction",
      Component == "V_d_E_In_only" ~ "Interception",
      Component == "SettlingVelocity_ms" ~ "Settling",
      Component == "V_d_ms" ~ "Total"
    )
  )
```

\newpage

### Parameter table

```{r Emerson fig 2 Pars}

TableData <- Output %>%
  select(
    Parametrization, LUC, Season,
    T_air_K, AirPressure_Pa, SurfaceIsWet_bool,
    SurfaceIsVegetated, CharacteristicRadius_m,
    ImpactionParameterAlpha, ParticleDensity_kgm3,
    BrownianDiffusionParameterGamma,
    R_a_sm
  ) %>%
  distinct() %>%
  rename(
    Gamma = BrownianDiffusionParameterGamma
  ) %>%
    mutate(
      Gamma = round(Gamma,3),
    C_b = GetParameters(Parametrization = Parametrization, TargetParameter = "C_b"),
    nu = GetParameters(Parametrization = Parametrization, TargetParameter = "nu"),
    C_In = GetParameters(Parametrization = Parametrization, TargetParameter = "C_In"),
    beta = GetParameters(Parametrization = Parametrization, TargetParameter = "beta"),
    C_Im = GetParameters(Parametrization = Parametrization, TargetParameter = "C_Im")
  ) %>%
  pivot_longer(
    cols = -c("Parametrization", "LUC"),
    values_transform = as.character,
    names_to = "Parameter"
  ) %>%
  mutate(
    Parametrization_LUC = paste0(Parametrization, "_LUC", LUC)
  ) %>%
  select(-Parametrization, -LUC) %>%
  pivot_wider(
    names_from = "Parametrization_LUC",
    values_from = "value"
  )

kable(
  x = TableData
)
```

### Plots

```{r Emerson fig 2B Plots, fig.width=8, fig.height=5}
# Load data extracted from Emerson et al. (2020)
ValidationData <- dd_subprocess_validation %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("Zhang01", "Emerson20")
    ),
    Type = "Extracted from Emerson et al. (2020)"
  )
ProcessContributions <- ProcessContributions %>%
  mutate(
    Label = factor(
      x = Parametrization,
      levels = c("Zhang01", "Emerson20")
    ),
    Type = "Calculated with ddpart"
  )

Colors <- c("black", "blue", "orange", "darkred", "purple")
names(Colors) <- c("Total", "Brownian", "Settling", "Interception", "Impaction")
Shapes <- c(3)
names(Shapes) <- c("Calculated with ddpart")
LineTypes <- c("solid")
names(LineTypes) <- c("Extracted from Emerson et al. (2020)")

ggplot() +
  geom_line(
    data = ValidationData,
    mapping = aes(
      x = ParticleDiameter_um,
      y = DepositionVelocity_cms,
      color = Process,
      linetype = Type
    )
  ) +
  geom_point(
    data = ProcessContributions,
    mapping = aes(
      x = ParticleDiameter_m * 1e6,
      y = v * 100,
      color = ComponentName,
      shape = Type
    ),
    size = 2
  ) +
  scale_shape_manual(values = Shapes) +
  scale_color_manual(values = Colors) +
  scale_linetype_manual(values = LineTypes) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3, 1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.box = "vertical",
    panel.spacing = unit(1, "lines")
    # legend.margin=margin()
  ) +
  facet_wrap(~Label)
```
Figure 2: Contribution of Brownian diffusion, Gravitational settling, Interception and Impaction to the total dry deposition velocity according to the Zhang et al. (2001) parametrization (left panel) and the Emerson et al. (2020) parameterization (right panel). Lines indicate data extracted from fig. 2 in Emerson et al. (2020). Results from the ddpart R package are indicated by "+". The figures show results at 0.4 m/s friction velocity for needleleaf forest (LUC 1) and a particle density of 1500 kg/m3. Data from ddpart is generated for season 1 with aerodynamic resistance set to zero, no hygroscopic swelling and dry surface.



\newpage

# References

 - Seinfeld JH, Pandis SN. Atmospheric Chemistry and Physics: From Air Pollution to Climate Change. Second edition. Wiley; 2006.

 - Zhang L, Gong S, Padro J, Barrie L. A size-segregated particle dry deposition scheme for an atmospheric aerosol module. Atmospheric Environment 2001;35:549–560.

 - Emerson EW, Hodshire AL, DeBolt HM, Bilsback KR, Pierce JR, McMeeking GR, Farmer DK. Revisiting particle dry deposition and its role in radiative effect estimates. Proceedings of the National Academy of Sciences 2020;117:26076–26082.
