---
title: "ddpart"
author: "Andreas Schmitz"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{ddpart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Package installation

Install the package
```{r install}
devtools::install_github(
  repo = "https://github.com/AndSchmitz/ddpart"
)
library(ddpart)
```

Dependencies for this vignette
```{r deps}
# library(tidyverse)
library(knitr)
# library(devtools)
library(scales)
library(ggplot2)
```

# Introduction

blending height concept


# Exampe 1: Vertical wind speed profile

## A) Comparison to power-law approach

The following code calculates the wind speed profile for two example cases (stable and unstable conditions) according to two approaches:
 1. According to Monin-Obukhov similarity theory, as generally used in this package. The implementation follows Seinfeld and Pandis (2006) chapter 16.4.3.
 2. For comparison, also the simpler power-law equation for the wind speed profile is shown, according to Seinfeld and Pandis (2006) chapter 16.4.5.

### Set parameters

The following code defines the parameters for the setup. Roughness length, radiation, wind speed, etc. are chosen to yield two specific values of the Monin-Obukhov lenght (L=-10 (relatively unstable conditions), L = 100 (relatively stable conditions)). The parameter *WindSpeedBlendingHeight_m* is set to several different values between 5 and 50 m. 

```{r WindSpeedProfile_SetParameters}

nHeightIntervals <- 20
MaxHeight_m <- 50

#Define input for unstable conditions
InputTable_UnstableConditions <- data.frame(
  Label = "Relatively unstable conditions",
  RoughnessLengthAnemometer_m = 0.01,
  ZeroPlaneDisplacementHeightAnemometer_m = 7 * 0.01,
  WindSpeedAtAnemometerHeight_ms = 3,
  AnemometerHeight_m = 10,
  SunAngle_degree =  20,
  T_air_K = 273 + 20,
  RelHum_percent = 0, #Does not matter if AerosolType = "Dry",
  AirPressure_Pa = 1013.25 * 100,
  GlobalRadiation_W_m2 = 500,
  CloudCover_percent = 10,
  Season = 1, #midsummer
  WindSpeedBlendingHeight_m = seq(5,MaxHeight_m,length.out = nHeightIntervals),
  TargetLUCCodeZhang2001 = 6, #grass
  #The following parameters do not matter for this use case:
  SurfaceIsWet_bool = F,
  ReferenceHeight_m = 2.5,
  RoughnessLengthTargetLUC_m = 0.03,
  ZeroPlaneDisplacementHeightTargetLUC_m = 0.21,
  #Particle properties
  DryParticleDiameter_m =  4.5 * 1e-6,
  ParticleDensity_kgm3 = 2 / 1e3 * 1e4,
  AerosolType = "Dry"
)

#Define input for stable conditions
InputTable_StableConditions <- InputTable_UnstableConditions %>%
  mutate(
    Label = "Relatively stable conditions",
    WindSpeedAtAnemometerHeight_ms = 1,
    RoughnessLengthAnemometer_m = 0.46,
    ZeroPlaneDisplacementHeightAnemometer_m = 7 * RoughnessLengthAnemometer_m,
    SunAngle_degree = -20,
    CloudCover_percent = 80
  )

#Combine inputs
InputTable <- bind_rows(
  InputTable_UnstableConditions,
  InputTable_StableConditions
)

```

### Calculate wind speed profile according to Monin-Obukhov theory

For each row of the *InputTable*, the wrapper function *CalculateDepositionVelocity()* is called and the wind speed at the respective blending height (among other parameters) is calculated.

```{r WindSpeedProfile_CallFunction, echo=FALSE}

#Calculate wind speed at blending height (and other aspects like dry deposition velocity)
Results <- CalculateDepositionVelocity(
  InputTable = InputTable
)

TableDat <- Results %>%
  select(Label, RoughnessLengthAnemometer_m, WindSpeedAtAnemometerHeight_ms, PasquillClass, ObukhovLength_Anemometer_m) %>%
  distinct()
kable(
  x = TableDat
)
```

### Calculate wind speed profile according to power-law approach

Seinfeld and Pandis (2006) chapter 16.4.5 empirical equation for the mean wind speed:

$u(TargetHeight) = u(AnemometerHeight) * (TargetHeight / AnemometerHeight)^p$

Exponent p taken from Seinfeld and Pandis (2006) figure 16.9 a):

- p = 0.087 for z0 = 0.01 and L=-10 (relatively unstable conditions)
- p = 0.43 for z0 = 0.46 and L = 100 (relatively stable conditions)

```{r WindSpeedProfile_AddPowerLawApproach}
Results <- Results %>%
  mutate(
    PowerLawExponent = case_when(
      Label == "Relatively unstable conditions" ~ 0.087,
      Label == "Relatively stable conditions" ~ 0.43,
      T ~ 999
    ),
    WindSpeedAtBlendingHeight_ms_PowerLaw = WindSpeedAtAnemometerHeight_ms * (WindSpeedBlendingHeight_m / AnemometerHeight_m)^PowerLawExponent
  ) %>%
  rename(
    WS_MO = WindSpeedAtBlendingHeight_ms,
    WS_PowerLaw = WindSpeedAtBlendingHeight_ms_PowerLaw
  )

PlotData <- Results %>%
  select(Label, WindSpeedBlendingHeight_m, WS_MO, WS_PowerLaw) %>%
  pivot_longer(
    cols = c("WS_MO","WS_PowerLaw"),
    values_to = "WS",
    names_to = "WS_type"
  ) %>%
  mutate(
    Approach = case_when(
      WS_type == "WS_PowerLaw" ~ "Power law approximation",
      WS_type == "WS_MO" ~ "Monin-Obukhov theory",
      T ~ "UNDEF"
    )
  )
```

### Results

The power law approach and the approach according to Monin-Obukhov theory yield similar vertical wind speed profiles. Differences likely result from the fact that the power law approach is only an approximation to the more accurate Monin-Obukhov theory approach.

```{r WindSpeedProfile_Plot, fig.width = 7, fig.height = 5}

ggplot(
  data = PlotData,
  mapping = aes(
    x = WS,
    y = WindSpeedBlendingHeight_m,
    color = Approach,
    shape = Label,
    group = interaction(Label, Approach)
  )
) +
geom_point() +
geom_line() +
ylim(0,max(Results$WindSpeedBlendingHeight_m)) +
xlab("Wind speed (m/s)") +
ylab("Height (m)") +
guides(
  color = guide_legend(
    title = ""
  ),
  shape = guide_legend(
    title = ""
  )
) +
theme(
  legend.position = "bottom",
  legend.box="vertical"
)
```

## B) Comparison to an example in Seinfeld and Pandis (2006)

Seinfeld and Panis (2006) page 745 provide the following example:

 - Known:
    - z0 = 0.0015 m
    - Anemometer height = 4 m
    - Wind speed = 7.8 m/s
    - Neutral atmopsheric stability
 - Derived:
    - Friction velocity = 0.4 m/s
    - Wind speed in 0.5 m = 5.8 m/s
    - Wind speed in 16 m = 9.3 m/s

This is reproduced in the following setup. Meteorological parameters are set up to yield Pasquill class D (neutral stability).

```{r WindSpeedExample}
InputTable <- data.frame(
  RoughnessLengthAnemometer_m = 0.0015,
  AnemometerHeight_m = 4,
  WindSpeedAtAnemometerHeight_ms = 7.8,
  WindSpeedBlendingHeight_m = c(0.5, 16),
  #
  ZeroPlaneDisplacementHeightAnemometer_m = 0.0015*7,
  SunAngle_degree =  20,
  T_air_K = 273+20,
  RelHum_percent = 0, #Does not matter if AerosolType = "Dry",
  AirPressure_Pa = 1013.25 * 100,
  GlobalRadiation_W_m2 = 500,
  CloudCover_percent = 10,
  Season = 1, #midsummer
  TargetLUCCodeZhang2001 = 6, #grass
  #Irrelevant:
  SurfaceIsWet_bool = F,
  ReferenceHeight_m = 2.5,
  RoughnessLengthTargetLUC_m = 0.03,
  ZeroPlaneDisplacementHeightTargetLUC_m = 0.21,
  DryParticleDiameter_m =  4.5 * 1e-6,
  ParticleDensity_kgm3 = 2 / 1e3 * 1e4,
  AerosolType = "Dry"
)

Results <- CalculateDepositionVelocity(
  InputTable = InputTable
)

kable(
  x = Results %>%
    select(PasquillClass, WindSpeedBlendingHeight_m, FrictionVelocity_Anemometer_ms, WindSpeedAtBlendingHeight_ms)
)

```

# Example 2: Dry deposition velocities

Emerson et al. (2020) figures 1 and 2 show the dry deposition velocity as a function of particle diameter for different settings (i.e. for different land use types, friction velocity, etc.). In the following, these figures are reproduced as a validation of the functions in the ddpart package.

## A) Contributions from collection efficiencies

Emerson et al. (2020) figure 2B shows the contribution of the four dry deposition sub-processes (Brownian motion, interception, impaction and gravitational settling) to the total dry deposition velocity. This figure is reconstructed with functions from the ddpart package. 

### Definition of parameters

```{r Emerson fig 2B parameters}

#Vector of particle diameters
d_p_vec <- 10^seq(-2,2, length.out = 100)
tmp <- expand.grid(1:9,d_p_vec)
d_p_vec <- tmp[,1] * tmp[,2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec  <- d_p_vec * 1e-6


InputTable_EmersonFig2B <- data.frame(
  Parametrization = "Emerson20",
  #These parameters are documented in Emerson et al. (2020)
  TargetLUCCodeZhang2001 = 1, #Evergreen needle leaf
  DryParticleDiameter_m =  d_p_vec,
  ParticleDensity_kgm3 = 1500,
  #These parameters are assumed
  AerosolType = "Dry",
  SurfaceIsWet_bool = F,
  #Parameter reference height does not matter
  ReferenceHeight_m = 20,
  #These parameters are tweaked to yield a friction velocity of approx 0.4 m/s.
  WindSpeedBlendingHeight_m = 50,
  RoughnessLengthAnemometer_m = 0.03,
  ZeroPlaneDisplacementHeightAnemometer_m = 7 * 0.03,
  WindSpeedAtAnemometerHeight_ms = 2.25,
  AnemometerHeight_m = 10,
  SunAngle_degree =  40,
  T_air_K = 273 + 20,
  RelHum_percent = 0, #Does not matter if AerosolType = "Dry",
  AirPressure_Pa = 1013.25 * 100,
  GlobalRadiation_W_m2 = 500,
  CloudCover_percent = 50,
  Season = 1 #midsummer
) %>%
  mutate(
    #These parameters do not matter because wind speed is tweak to reproduce friction velocity = 0.4 m/s
    #z0 corresponding to LUC
    RoughnessLengthTargetLUC_m = GetLandUseParameters(
      LUCs = TargetLUCCodeZhang2001,
      Seasons = Season,
      TargetPar = "z_0_m",
      Parametrization = Parametrization
    ),
    #0.78*20m = 15.6 for forest as in the EMEP model
    ZeroPlaneDisplacementHeightTargetLUC_m = 15.6
  )
```

### Calculations
```{r Emerso fig 2B calculations}

#Calculate vd and other outputs
Results <- CalculateDepositionVelocity(
  InputTable = InputTable_EmersonFig2B,
  Parametrization = "Emerson20"
)

#Print resulting friction velocity
#Is approx. 0.4 as in Emerson et al. (2020) figure 2B
print(unique(Results$FrictionVelocityTargetLUC_ms))

#Plot separate dry deposition processes contributions
ProcessContribution <- Results %>%
  mutate(
    BounceCorrectionTerm = ifelse(
      test = SurfaceIsWet_bool,
      yes = 1,
      no = exp(-sqrt(StokesNumber))
    ),
    #Calculation of resistances and corresponding vd's for each process assuming contributions from other
    #processes are zero.
    epsilon_0 = GetParameters(TargetParameter = "epsilon_0", Parametrization = Parametrization),
    R_s_E_b_only = 1 / (epsilon_0 * FrictionVelocityTargetLUC_ms * BounceCorrectionTerm * E_b),
    V_d_E_b_only = 1 / (R_a_sm + R_s_E_b_only),
    R_s_E_Im_only = 1 / (epsilon_0 * FrictionVelocityTargetLUC_ms * BounceCorrectionTerm * E_Im),
    V_d_E_Im_only = 1 / (R_a_sm + R_s_E_Im_only),
    R_s_E_In_only = 1 / (epsilon_0 * FrictionVelocityTargetLUC_ms * BounceCorrectionTerm * E_In),
    V_d_E_In_only = 1 / (R_a_sm + R_s_E_In_only)
    #Sanity checked V_d_sum against V_d -> ok.
    # V_d_sum = V_d_E_b_only + V_d_E_Im_only + V_d_E_In_only + V_g
  ) %>%
  select(DryParticleDiameter_m,V_d_E_b_only,V_d_E_Im_only,V_d_E_In_only,V_d_RefHeight_ms, SettlingVelocity_ms) %>%
  pivot_longer(
    cols = -c("DryParticleDiameter_m"),
    names_to = "Component",
    values_to = "v"
  ) %>%
  mutate(
    ComponentName = case_when(
      Component == "V_d_E_b_only" ~ "Brownian",
      Component == "V_d_E_Im_only" ~ "Impaction",
      Component == "V_d_E_In_only" ~ "Interception",
      Component == "SettlingVelocity_ms" ~ "Gravitational settling",
      Component == "V_d_RefHeight_ms" ~ "Total dry deposition"
    )
  )
```


### Results

The following figure reproduces Emerson et al. (2020) figure 2B.

```{r Emerso fig 2B plotting, fig.width = 7, fig.height = 5, caption = "Test"}
Colors <- c("black","blue","orange","darkred","purple")
names(Colors) <- c("Total dry deposition","Brownian","Gravitational settling","Interception","Impaction")
ggplot(
  data = ProcessContribution %>%
    filter(
      # Component == "V_d_E_b_only"
    ),
  mapping = aes(
    x = DryParticleDiameter_m*1e6,
    y = v*100,
    color = ComponentName
  )
) +
  scale_color_manual(values = Colors) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3,1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  geom_line(
    size = 2
  ) +
  # ylim(-3,2) +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )

```


## B) Deposition velocities for different LUCs

Emerson et al. (2020) figure 1 shows the dry deposition velocity for different land use classes. This figure is reconstructed with functions from the ddpart package. 

### Definition of parameters

```{r Emerson fig 1 parameters}

#Vector of particle diameters
d_p_vec <- 10^seq(-2,2, length.out = 100)
tmp <- expand.grid(1:9,d_p_vec)
d_p_vec <- tmp[,1] * tmp[,2]
d_p_vec <- d_p_vec[d_p_vec <= 100]
d_p_vec  <- d_p_vec * 1e-6


InputTable_EmersonFig1 <- data.frame(
  Parametrization = "Emerson20",
  #These parameters are documented in Emerson et al. (2020)
  TargetLUCCodeZhang2001 = 1, #Evergreen needle leaf
  DryParticleDiameter_m =  d_p_vec,
  ParticleDensity_kgm3 = 1200,
  #These parameters are assumed
  SurfaceIsWet_bool = F,
  AerosolType = "Dry",
  RelHum_percent = 80,
  #These parameters are tweaked to yield a friction velocity of approx 0.2 m/s.
  ReferenceHeight_m = 20,
  WindSpeedBlendingHeight_m = 50,
  RoughnessLengthAnemometer_m = 0.03,
  ZeroPlaneDisplacementHeightAnemometer_m = 7 * 0.03,
  WindSpeedAtAnemometerHeight_ms = 0.95,
  AnemometerHeight_m = 10,
  SunAngle_degree =  40,
  T_air_K = 273 + 20,
  AirPressure_Pa = 1013.25 * 100,
  GlobalRadiation_W_m2 = 500,
  CloudCover_percent = 50,
  Season = 1 #midsummer
) %>%
  mutate(
    #These parameters do not matter because wind speed is tweak to reproduce friction velocity
    #z0 corresponding to LUC
    RoughnessLengthTargetLUC_m = GetLandUseParameters(
      LUCs = TargetLUCCodeZhang2001,
      Seasons = Season,
      TargetPar = "z_0_m",
      Parametrization = Parametrization
    ),
    #0.78*20m = 15.6 for forest as in the EMEP model
    ZeroPlaneDisplacementHeightTargetLUC_m = 15.6
  )

# tmp1 <- InputTable_EmersonFig1 %>%
#   mutate(
#     ReferenceHeight_m = 100
#     #SurfaceIsWet_bool = F
#   )
# 
# tmp2 <- InputTable_EmersonFig1 %>%
#   mutate(
#     WindSpeedAtAnemometerHeight_ms = 4
#   )
# 
# InputTable_EmersonFig1 <- bind_rows(
#   InputTable_EmersonFig1,
#   tmp1
#   # tmp2
# )
```


### Calculations
```{r Emerso fig 1 calculations,  fig.width = 7, fig.height = 5, caption = "Test"}

#Calculate vd and other outputs
Results_EmersonFig1 <- CalculateDepositionVelocity(
  InputTable = InputTable_EmersonFig1,
  Parametrization = Parametrization
)


#Print resulting friction velocity
#Is approx. 0.2 as in Emerson et al. (2020) figure 1
print(unique(Results_EmersonFig1$FrictionVelocityTargetLUC_ms))

 

ggplot(
  data = Results_EmersonFig1,
  mapping = aes(
    x = DryParticleDiameter_m*1e6,
    y = V_d_RefHeight_ms*100,
    color = as.character(SurfaceIsWet_bool)
  )
) +
  # scale_color_manual(values = Colors) +
  scale_x_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(1e-3,1e2),
    expand = expansion(mult = c(0, 0))
  ) +
  annotation_logticks() +
  geom_line(
    size = 2
  ) +
  # ylim(-3,2) +
  xlab("Particle diameter (um)") +
  ylab("Deposition velocity (cm/s)") +
  theme(
    legend.position = "bottom"
    # legend.title = element_blank()
  )


tmp <- Results_EmersonFig1 %>%
  filter(
    ParticleDiameter_m == min(ParticleDiameter_m)
  ) %>%
  select(ParticleDiameter_m, FrictionVelocityTargetLUC_ms, E_b, E_In, E_Im, R_a_sm, R_s_sm) %>%
  distinct()
print(tmp)
```


# Example 3: Hygroscopic swelling

### Definition of parameters

```{r Hygroscopic swelling,  fig.width = 7, fig.height = 5, caption = "Test"}

DryParticleDiameter_m <- c(0.1, 0.5, 1.5, 4.5) * 1e-6
AerosolType <- c("Dry", "SeaSalt", "Urban", "Rural", "AmmoniumSulfate")
RelHum_percent <- seq(0, 100, by=5)
ExampleData <- expand_grid(DryParticleDiameter_m, AerosolType, RelHum_percent) %>%
  mutate(
    DryParticleDiameter_m = DryParticleDiameter_m,
    WetParticleDiameter_m = CalculateHygroscopicSwelling(
      DryParticleDiameter_m = DryParticleDiameter_m,
      AerosolType = AerosolType,
      RelHum_percent = RelHum_percent
    ),
    RelChangeParticleDiameter_percent = round( (WetParticleDiameter_m - DryParticleDiameter_m) / DryParticleDiameter_m * 100, 2),
    FacetLabel = paste("Dry diameter =", DryParticleDiameter_m /  1e-6, "um")
  )

ggplot(
  data = ExampleData,
  mapping = aes(
    x = RelHum_percent,
    y = WetParticleDiameter_m,
    color = AerosolType
  )
) +
geom_line() +
scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) +
facet_wrap( ~ FacetLabel, scales = "free_y") +
xlab("Relative humidity (%)") +
ylab("Wet particle diameter (um)") +
guides(
  color = guide_legend(title = "")
) +
theme(
  legend.position = "bottom"
)

ggplot(
  data = ExampleData,
  mapping = aes(
    x = RelHum_percent,
    y = RelChangeParticleDiameter_percent,
    color = AerosolType
  )
) +
geom_line() +
# scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) +
facet_wrap(
  facets = ~ FacetLabel
  # scales = "free_y"
) +
xlab("Relative humidity (%)") +
ylab("Relative change in particle diameter (%)") +
guides(
  color = guide_legend(title = "")
) +
theme(
  legend.position = "bottom"
)
```


# References

Seinfeld JH, Pandis SN. Atmospheric Chemistry and Physics: From Air Pollution to Climate Change. Second edition. Wiley; 2006.

Zhang L, Gong S, Padro J, Barrie L. A size-segregated particle dry deposition scheme for an atmospheric aerosol module. Atmospheric Environment 2001;35:549–560.

Emerson EW, Hodshire AL, DeBolt HM, Bilsback KR, Pierce JR, McMeeking GR, Farmer DK. Revisiting particle dry deposition and its role in radiative effect estimates. Proceedings of the National Academy of Sciences 2020;117:26076–26082.


